---
title: "Volcanoes"
author: Laura Brugger
output: html_document
---

```{r setup, include=FALSE}
#knitr::opts_chunk$set(echo = TRUE)
```

The following packages are required. If not already installed, do so using `install.packages`: 
`ggplot2`
`dplyr`
`readxl`

##Visualizing Elevation of Volcanoes with Persistent Lava Lakes

First, let's practice making a data frame. This data frame will include 7 volcanoes with persistent lava lakes plus the tallest volcano in the world, their location, their elevation, and their most recent known eruption. 

```{r}
volcano <- c('Erta Ale', 'Ambrym', 'Mount Yasur', 'Mount Erebus', 'Kilauea', 'Nyiragongo', 'Masaya', 'Mauna Loa')
location <- c('Ethiopia','Vanuatu','Vanuatu','Antarctica','United States', 'Democratic Republic of the Congo','Nicaragua', 'United States')
elevation <- c(2011, 4377, 1184, 12448, 4091, 11385, 2083, 13678)
eruption <- c('2009', '2019', '2016', '2011', '2018', '2016', '2008', '1984')

volcData <- data.frame(volcano, location, elevation, eruption)
```

Create a bar chart to display the elevation of each volcano. 

```{r}
library(ggplot2)
ggplot(data=volcData, aes(x=volcano, y=elevation, fill=volcano)) +
  geom_bar(stat="identity") + 
  ggtitle("Elevation of Volcanoes with Active Lava Lakes", subtitle="An R-Ladies Masterpiece") + 
  ylab("Elevation (in feet)") + 
  xlab ("Volcano") + 
  theme(legend.position = "none")

```

But Wait! Mauna Loa doesn't have an active lava lake. Let's get rid of it. Before we do that, this is a good time to save, add, commit, and push our work to GitHub. Go to your terminal and enter the following: 

```{r}
#git add -A
#git commit -a -m "add initial lava lake bar chart" 
  #Remember, our commit message is a breadcrumb for us to remember why we made this commit
#git push origin master 

```

Use `filter` from the `dplyr` package to remove the volcano "Mauna Loa" and assign to a new dataframe called `lavaLake`. 

```{r}
library(dplyr)
volcData %>%
  filter (volcano != "Mauna Loa") -> lavaData
```

Let's try again! Make a new bar chart with our new `lavaLake` data. 
Let's also reorder from highest to lowest elevation. 

```{r}
ggplot(data=lavaData, aes(x= reorder (volcano, -elevation), y=elevation, fill=volcano)) +
  geom_bar(stat="identity") + 
  ggtitle("Elevation of Volcanoes with Active Lava Lakes", subtitle="An R-Ladies Masterpiece") + 
  ylab("Elevation (in feet)") + 
  xlab ("Volcano") + 
  theme(legend.position="none")
```

Save, add, commit, push! 

```{r}
#git add -A
#git -a -m "add revised lava lake bar chart that omits Mauna Loa" 
#git push origin master
```

Try it on your own and make a graph of the last known eruption for each volcano. If you'd like, go ahead and save, add, commit, and push changes. 

```{r}
ggplot(data= , aes(x= , y= , fill= )) +
  geom_bar(stat = "identity")

```

#Bringing in external data

Using `readxl`, bring in the dataset included in the repo we downloaded from GitHub.This includes the 5 most recent eruptions of select volcanoes.  

```{r}
library(readxl)
eruptionData <- read_xlsx("volcano.xlsx")
```

For practice, let's get rid of Mauna Loa again so we only graph volcanic eruptions of those with persistent lava lakes. 

```{r}
eruptionData %>%
  filter (volcano != "Mauna Loa") -> lavaEruptData
```


Using `ggplot`, make a scatterplot that displays the 5 most recent eruptions of volcanoes with persistent lava lakes. Give your graph a title, a subtitle, and add labels to the x and y axes. 
```{r}
ggplot(data=lavaEruptData, aes(y=, x=, color= volcano, position="dodge")) +
  geom_point() + 
  ggtitle("", subtitle="") + 
  ylab("") + 
  xlab ("") + 
  theme(legend.position="none")
  
```

You know the drill: Save, add, commit, push! 

```{r}
#git -A
#git -a -m "added new section to include recent eruption scatterplot" 
#git commit push origin
```