---
title: "Volcanoes"
author: Laura Brugger
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The following packages are required. Install using `install.packages`
`ggplot2`
`dplyr`
`purrr`

##Visualizing Elevation of Volcanoes with Persistent Lava Lakes

First, let's practice making a data frame. This data frame will include 7 volcanoes with persistent lava lakes plus the tallest volcano in the world, their location, their elevation, and their most recent known eruption. 

```{r}
volcano <- c('Erta Ale', 'Ambrym', 'Mount Yasur', 'Mount Erebus', 'Kilauea', 'Nyiragongo', 'Masaya', 'Mauna Loa')
location <- c('Ethiopia','Vanuatu','Vanuatu','Antarctica','United States', 'Democratic Republic of the Congo','Nicaragua', 'United States')
elevationFt <- c(2011, 4377, 1184, 12448, 4091, 11385, 2083, 13678)
lastKnownEruption <- c('2009', '2019', '2016', '2011', '2018', '2016', '2008', '1984')

volcanoData <- data.frame(volcano, location, elevationFt, lastKnownEruption)
```

Create a bar chart to display the elevation of each volcano. 

```{r}
library(ggplot2)
ggplot(data=volcanoData, aes(x=volcano, y=elevationFt, fill=volcano)) +
  geom_bar(stat="identity") + 
  ggtitle("Elevation of Volcanoes with Active Lava Lakes", subtitle="An R-Ladies Masterpiece") + 
  ylab("Elevation (in feet)") + 
  xlab ("Volcano") + 
  theme(legend.position = "none")

```

But Wait! Mauna Loa does NOT have an active lava lake. Let's get rid of it. Before we do that, this is a good time to save, add, commit, and push our work to GitHub. Go to your terminal and enter the following: 

```{r}
#git -A
#git -a -m "add initial lava lake bar chart" 
  #Remember, our commit message is a breadcrumb for us to remember why we made this commit
#git commit push origin

```

Use `filter` from the `dplyr` package to remove the volcano "Mauna Loa" and assign to a new dataframe called `lavaLake`. 

```{r}
library(dplyr)
volcanoData %>%
  filter (volcano != "Mauna Loa") -> lavaLake
```

Let's try again! Make a new bar chart with our new `lavaLake` data. 
Let's also reorder from highest to lowest elevation. 

```{r}
ggplot(data=lavaLake, aes(x= reorder (volcano, -elevationFt), y=elevationFt, fill=volcano)) +
  geom_bar(stat="identity") + 
  ggtitle("Elevation of Volcanoes with Active Lava Lakes", subtitle="An R-Ladies Masterpiece") + 
  ylab("Elevation (in feet)") + 
  xlab ("Volcano") + 
  theme(legend.position="none")
```

Save, add, commit, push! 

```{r}
#git -A
#git -a -m "add revised lava lake bar chart that omits Mauna Loa" 
#git commit push origin
```

Try it on your own and make a graph of the last known eruption fro each volcano. 

```{r}
ggplot(data=lavaLake, aes(x=volcano, y=lastKnownEruption, fill= volcano)) +
  geom_bar(stat = "identity")

```

```{r}
library(readxl)
eruptions <- read_xlsx("volcano.xlsx")
```

Let's get rid of Mauna Loa again. 
```{r}
eruptions %>%
  filter (volcano != "Mauna Loa") -> lavaEruption
```

```{r}
ggplot(data=lavaEruption, aes(y=volcano, x=eruption, color= volcano, position="dodge")) +
  geom_point() + 
  ggtitle("Recent Eruptions of Volcanoes with Persistent Lava Lakes", subtitle="An R-Ladies Masterpiece") + 
  ylab("Volcano") + 
  xlab ("Year of Eruption") + 
  theme(legend.position="none")
  
```

You know the drill: Save, add, commit, push! 

```{r}
#git -A
#git -a -m "added new section to include recent eruption scatterplot" 
#git commit push origin
```